A peek into the inner workings of pandemic prediction models

### 2020-10-30 text/20/sir-model.Rmd
More work on sir paper

```{r, echo=FALSE}
suppressMessages(suppressWarnings(library(knitr)))
suppressMessages(suppressWarnings(library(tidyverse)))
knitr::opts_chunk$set(echo=FALSE, fig.width=6, fig.height=3)
```

@@

@@

@@

>>>>

```{r, echo=FALSE}
suppressMessages(suppressWarnings(library(knitr)))
suppressMessages(suppressWarnings(library(tidyverse)))
suppressMessages(suppressWarnings(library(usmap)))

knitr::opts_chunk$set(echo=FALSE, fig.width=6, fig.height=3)
```

@@

### A vaccination model

When you get a vaccine, the dynamics change. The susceptible

@@

### Agent based models

An agent based model replaces buckets of water with individual particles. There particles move around at varying speeds and when an infected particle collides a susceptible particle, the latter also becomes infected. Think of it as a giant simulated game of tag, with the infected people playing the role of "it". You can control each particle differently. Faster moving particles represent individuals who do not practice social distancing well. You can assign each particle a different level of susceptibility as well, so that some become infected with very high probability after a collision and others become infected with a much lower probability. These models can include other sources of variation, such as changes in social distancing after a lockdown.

Agent based models have been compared to a giant game of the Sims. The number of people in the agent based models, however, is often in the thousands or even millions.

### Hierarchical models

A hierarchical model divides up the country into regions (typically individual counties) and lets each region have a different SIR model. With 3,141 counties, this represents a huge increase in complexity, one that could potentially choke even the fastest computers. The saving grace in these hierarchical models is that you can often rely on the fact that similar regions will behave similarly in the dynamics of the SIR model.

You can model the similarity of regions using geographic proximity (New York counties and New Jersey counties should have similar dynamics, for example). You could also model the similarities using U.S. Census information (counties with very low population densities should have similar dynamics, for example).

### Time varying parameters

Time is one of the biggest sources of heterogeneity. The spread of an infectious disease changes over time as personal behavior changes and as government interventions change.

@@

https://www.nature.com/articles/d41586-020-01003-6
